---
title: ivDiag usage examples
author: Apoorva Lal
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    self_contained: true
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r, include = F}
library(knitr); library(ggplot2)
knitr::opts_chunk$set(echo = T, include = T, warning = F, message = F, cache   = T)
theme_set(theme_minimal())
```


`ivDiag` is a package for uncertainty quantification and sensitivity
analysis for IV designs that provides `R` implementations of the
guidelines proposed in Lal et al (2021).

The package provides functions to compute (block-)bootstrapped
standard errors, F-statistics, and weak-IV robust confidence intervals
as well as the Local-to-Zero estimator (Conley, Hansen, and Ross
2012). Estimation routines in `ivDiag` always return lists that can be sliced and appended to additional rows in tables produced by `modelsummary`, `fixest::etable` and `stargazer` etc.

This package is a work in progress. Github issues and/or pull requests are welcome.

# Uncertainty Quantification

```{r}
library(ivDiag); library(ggplot2)
# %% Meredith (2013) APSR data
df <-readRDS("../tmp/apsr_Meredith_2013.rds")
Y <- "DemShareDB"; D <- "DemShareGOV"; Z <- "HomeGOV"
controls <- c("HomeDB")
cl <- "fips"
FE <- c("fips","RaceID")
weights <- "Weight"
```


## First Stage Testing

```{r}
# %% first stage F statistics
fst = first_stage_tests(data=df, Y=Y, D=D, Z=Z, controls=controls, FE =FE,
  cl=cl, weights=weights, boot = T, nboot = 100)
fst |> print()
```


## Bootstrap / Block-bootstrap Standard Errors and F Statistic

```{r}
bootres=boot_IV(data=df, Y=Y, D=D, Z=Z, controls=controls, FE = FE,
  cl=cl, weights=weights, nboot = 100)
bootres |> print()
```


### Weak-IV robust Anderson-Rubin test

Fork of `ivmodel::AR.test` that uses `felm` to partial out and fit the
IV regression instead of `AER` and can fit larger models.

```{r}
AR_test(df, Y, D, Z, controls, FE) |> print()
```


# Sensitivity Analysis

## Zero-First-Stage

Example from guiso sapienza & zingales 2016.

```{r}
library(ivDiag); library(magrittr); library(patchwork); library(ggfortify); library(lfe)

source("../R/utils.R")
load('../tmp/GSZ2016.rds')

# formula
f = formula_lfe('totassoc_p',
  X = c('altitudine', 'escursione', 'costal', 'nearsea', 'population', 'pop2', 'gini_land', 'gini_income' ),
  W = 'libero_comune_allnord', Z = "bishopcity")

# vanilla clustered fit
m1 = felm(f, data = gsz_2016, weights = gsz_2016$population) |> robustify()

# bootstrap fit - this can also be done using `boot` with the corresponding ivreg / fixest call
m2 = felm(f, data = gsz_2016, weights = gsz_2016$population,
  Nboot = 1000, nostats=structure(FALSE,boot=TRUE))
iv_boots = m2$boot["`libero_comune_allnord(fit)`", ]
```

### LtZ estimates

```{r}
zfs_prior # estimate in ZFS south sample - saved in processing
ltz_est = ltz(gsz_2016_residualised, "W", "Z", mu = zfs_prior, sig = 0.1, ivmod = m1)
ltz_est |> print()
```

### Visualising analytic/bootstrap/ltz sampling distributions

`viz_iv_dists` expects `(iv_estimate, bootstrap_draws, ltz_estimate)` and
returns a list of ggplot objects with the corresponding sampling distributions.
These can then be wrapped into a single figure using `patchwork` or `gridExtra`.
Here, we use `patchwork::wrap_plots` to stack them.

```{r}
analytic_est = m1 %>% summary %>% .$coefficients %>% .[nrow(.), ] %>% .[1:2]
viz_iv_dists(analytic_est, iv_boots, ltz_est) %>% wrap_plots(nrow = 3)
```
